---
title: "TS Final Project_Reg Model(with ARIMA errors)"
author: "Louise"
date: "3/9/2022"
output: html_document
---

```{r setup, include=FALSE}
library("knitr")
library("rmarkdown")

library(tseries)
library(ggplot2)
library(forecast)
library(fpp)
library(TSA)
library(TSstudio)
#install.packages("tidyverse")
library(tidyverse)

library(zoo)
library(dplyr)

library(vars)
```

```{r}
library(readr)
urlfile="https://raw.githubusercontent.com/Beau-Smit/Time-Series/main/MN_Ag_YieldsAndInputs.csv"
dat <- read.csv(url(urlfile))
```

```{r}
#head(dat)
#tail(dat)
```
```{r}
# missing values
ag_df <- dat[!is.na(dat$Soy_BUperAcre),] %>% arrange(Year) 
#ag_df
```

- 3 useful variable: Corn_BUperAcre, Wheat_BUperAcre, Soy_BUperAcre, yearly data from 1934 to 2021
- each crop have 4 fertilizers, but too many nulls and gaps in those features ➔ will be ignored 

```{r}
# plot for relationship between variables
plot(ag_df$Corn_BUperAcre, ag_df$Wheat_BUperAcre, main = "Corn vs Wheat")
plot(ag_df$Corn_BUperAcre, ag_df$Soy_BUperAcre, main = "Corn vs Soy")
plot(ag_df$Soy_BUperAcre, ag_df$Wheat_BUperAcre, main = "Soy vs Wheat")
```

linear relationship

```{r}
# ts object
corn <- ts(ag_df$Corn_BUperAcre, frequency = 1, start = 1934)
wheat <- ts(ag_df$Wheat_BUperAcre, frequency = 1, start = 1934)
soy <- ts(ag_df$Soy_BUperAcre, frequency = 1, start = 1934)
```


```{r}
# corr
df <- cbind(corn, soy, wheat)
cor(df)
```


```{r}
# plot ts
autoplot(corn, main = "Crops Yield", xlab = "Year", ylab = "BU Per Acre", series = "Corn") +
  autolayer(wheat, series = "Wheat") +
  autolayer(soy, series = "Soy")
```

- boxcox needed
- potential seasonality or just trend? need to check

```{r}
# lambda calculation
BoxCox.lambda(corn)
BoxCox.lambda(soy)
BoxCox.lambda(wheat)
```

```{r}
# test with different lambda and plot ts to determine lambda for all three variables
corn_bc <- BoxCox(corn, lambda = 0.08)
soy_bc <- BoxCox(soy, lambda = 0.08)
wheat_bc <- BoxCox(wheat, lambda = 0.08)

autoplot(corn_bc, main = "Crops Yield after Boxcox", xlab = "Year", series = "Corn_bc") +
  autolayer(wheat_bc, series = "Wheat_bc") +
  autolayer(soy_bc, series = "Soy_bc")
```

can proceed with lambda = 0.5 (square root trans plus linear trans) 
or lambda = 0 (log trans) for easier interpretation

```{r}
# create variables for later use
corn_sqrt <- BoxCox(corn, lambda = 0.5)
soy_sqrt <- BoxCox(soy, lambda = 0.5)
wheat_sqrt <- BoxCox(wheat, lambda = 0.5)

corn_log <- BoxCox(corn, lambda = 0)
soy_log <- BoxCox(soy, lambda = 0)
wheat_log <- BoxCox(wheat, lambda = 0)

df_sqrt <- cbind(corn_sqrt, soy_sqrt, wheat_sqrt)
df_log <- cbind(corn_log, soy_log, wheat_log)

cor(df_sqrt) # check corr after boxcox
cor(df_log)
```


```{r}
# differencing and stationary check: determine d and D

diff(df_sqrt) %>% autoplot(main = "Crops Yield after sqrt+diff")
diff(df_log) %>% autoplot(maing = "Crops Yield after log+diff")

adf.test(diff(df_log)[,1])
adf.test(diff(df_sqrt)[,1])
```

both sqrt+1st differencing and log+1st differencing can bring to stationary ts, but log trans+1st differencing had relative constant variance.
- proceed with log + d=1 (no seasonal differencing)

```{r, include=FALSE}
# acf/pacf to determine potential p,q,P,Q
# eacf for a mix of AR and MA
head(df_log)
diff(diff(df_log)[,1]) %>% acf()
diff(diff(df_log)[,1]) %>% pacf()
# nsdiffs(diff(df_log)[,1]): Non seasonal data -> P,Q = 0?
```


```{r}
# train/test split
split <- ts_split(df, sample.out = 8)

train <- split$train
test <- split$test

head(train)
```

#### Regression
```{r}
# regression
tslm_corn_log <- tslm(formula = corn ~ soy + wheat,
                      data = train,lambda = 0, biasadj = TRUE)
tslm_wheat_log <- tslm(formula = wheat ~ corn + soy,
                       data = train,lambda = 0, biasadj = TRUE)
tslm_soy_log <- tslm(formula = soy ~ corn + wheat,
                     data = train, lambda = 0, biasadj = TRUE)
```


```{r}
# plot regression fitted values & actual values in training set
autoplot(train[,1], series = "corn") + 
  autolayer(tslm_corn_log$fitted.values, series = "fitted corn")

autoplot(train[,2], series = "soy") +
  autolayer(tslm_soy_log$fitted.values, series = "fitted soy")

autoplot(train[,3], series = "wheat") +
  autolayer(tslm_wheat_log$fitted.values, series = "fitted wheat")

```


```{r}
#summary(tslm_corn_log)
```


```{r}
checkresiduals(tslm_corn_log)
checkresiduals(tslm_wheat_log)
checkresiduals(tslm_soy_log)
```

Breusch-Godfrey test for all three regression models had a low p-value, indicating residuals were autocorrelated ➔use regression with ARIMA error model is appropriate

#### Reg model with ARIMA errors
```{r}
# model estimation
regARIMA_corn_log <- auto.arima(train[,1], xreg = train[,2:3], d = 1, 
                                 lambda = 0, biasadj = TRUE, 
                                trace = TRUE, ic = "aicc")
  
regARIMA_wheat_log <- auto.arima(train[,3], xreg = train[,1:2], d = 1,
                                  lambda = 0, biasadj = TRUE, 
                                 trace = TRUE, ic = "aicc")
  
regARIMA_soy_log <- auto.arima(train[,2], xreg = train[,c(1,3)], d = 1,
                                lambda = 0, biasadj = TRUE, 
                               trace = TRUE, ic = "aicc")
```


```{r}
summary(regARIMA_corn_log)
summary(regARIMA_soy_log)
summary(regARIMA_wheat_log)
```


```{r}
# plot fitted values based on regression with ARIMA errors model vs actual values in training set
autoplot(train[,1], series = "corn") + 
  autolayer(regARIMA_corn_log$fitted, series = "fitted corn")

autoplot(train[,2], series = "soy") +
  autolayer(regARIMA_soy_log$fitted, series = "fitted soy")

autoplot(train[,3], series = "wheat") +
  autolayer(regARIMA_wheat_log$fitted, series = "fitted wheat")
```


```{r}
# residauls(), checkresiduals(), zero mean, (const variance)
# autocorr test
# normality test

checkresiduals(regARIMA_corn_log$residuals)
Box.test(regARIMA_corn_log$residuals, type =  "Ljung-Box")
shapiro.test(regARIMA_corn_log$residuals)

checkresiduals(regARIMA_soy_log$residuals)
Box.test(regARIMA_soy_log$residuals, type =  "Ljung-Box")
shapiro.test(regARIMA_soy_log$residuals)

checkresiduals(regARIMA_wheat_log$residuals)
Box.test(regARIMA_wheat_log$residuals, type =  "Ljung-Box")
shapiro.test(regARIMA_wheat_log$residuals)
```

Ljung-Box test had large p-value(> 0.05) for all three models, showing these model had captured patterns in the series so that there were no autocorrelation in the residuals

Shapiro-Wilk test had large p-value(>0.05) for all three models as well, indicating the residuals are normally distributed

```{r}
# forecast
head(test)

forecast_corn <- forecast(regARIMA_corn_log, h = 8, level = 95, xreg = test[,2:3])
forecast_soy <- forecast(regARIMA_soy_log, h = 8, level = 95, xreg = test[,c(1,3)])
forecast_wheat <- forecast(regARIMA_wheat_log, h = 8, level = 95, xreg = test[,1:2])
```


```{r}
# forecast values from the model vs actual values test setting 
autoplot(train[,1], main = "Corn Yield Forecast", xlab = "BU Per Acre", ylab = "Year") + autolayer(test[,1], series = "corn") + 
  autolayer(forecast_corn$mean, series = "forecasted corn")

autoplot(train[,2], main = "Soy Yield Forecast", xlab = "BU Per Acre", ylab = "Year") + autolayer(test[,2], series = "soy") +
  autolayer(forecast_soy$mean, series = "forecasted soy")

autoplot(train[,3], main = "Wheat Yield Forecast", xlab = "BU Per Acre", ylab = "Year") + autolayer(test[,3], series = "wheat") + autolayer(forecast_wheat$mean, series = "forecasted wheat")
```

```{r}
# model evaluation (testing set)
accuracy(forecast_corn, test[,1])
accuracy(forecast_soy, test[,2])
accuracy(forecast_wheat, test[,3])
```



#### from baseline_models.Rmd but with testing set obs = 8
```{r}
# split train and test sets
first_test_yr <- 2014

corn_train <- window(corn, end=first_test_yr-1)
corn_test <- window(corn, start=first_test_yr)
wheat_train <- window(wheat, end=first_test_yr-1)
wheat_test <- window(wheat, start=first_test_yr)
soy_train <- window(soy, end=first_test_yr-1)
soy_test <- window(soy, start=first_test_yr)

h <- 8

# holt method
#plot(holt(corn_train, h=h, damped=T))
#plot(holt(wheat_train, h=h, damped=T, exponential=T))
#plot(holt(soy_train, h=h, damped=T))

# actual estimates
holt_predictions_corn <- holt(corn_train, h=h, damped=T)$mean
holt_predictions_soy <- holt(soy_train, h=h, damped=T, exponential=T)$mean
holt_predictions_wheat <- holt(wheat_train, h=h, damped=T)$mean

# auto.arima
lam_corn <- BoxCox.lambda(corn_train)
lam_soy <- BoxCox.lambda(soy_train)
lam_wheat <- BoxCox.lambda(wheat_train)

arima_mdl_corn <- auto.arima(corn_train, lambda = lam_corn)
arima_mdl_soy <- auto.arima(soy_train, lambda = lam_soy)
arima_mdl_wheat <- auto.arima(wheat_train, lambda = lam_wheat)

arima_predictions_corn <- (forecast(arima_mdl_corn, h=h))$mean
arima_predictions_soy <- (forecast(arima_mdl_soy, h=h))$mean
arima_predictions_wheat <- (forecast(arima_mdl_wheat, h=h))$mean

#plot(forecast(arima_mdl_corn, h=h))
#plot(forecast(arima_mdl_soy, h=h))
#plot(forecast(arima_mdl_wheat, h=h))
```

#### corn forecast accuracy 
```{r}
accuracy(forecast_corn, test[,1])
accuracy(holt_predictions_corn, test[,1])
accuracy(arima_predictions_corn, test[,1])
```

#### soy forecast accuracy 
```{r}
accuracy(forecast_soy, test[,2])
accuracy(holt_predictions_soy, test[,2])
accuracy(arima_predictions_soy, test[,2])
```

#### wheat forecast accuracy 
```{r}
accuracy(forecast_wheat, test[,3])
accuracy(holt_predictions_wheat, test[,3])
accuracy(arima_predictions_wheat, test[,3])
```

