---
title: "TS Final Project_Reg Model(with ARIMA errors)"
author: "Louise"
date: "3/9/2022"
output: html_document
---

```{r setup, include=FALSE}
library("knitr")
library("rmarkdown")

library(tseries)
library(ggplot2)
library(forecast)
library(fpp)
library(TSA)
library(TSstudio)
#install.packages("tidyverse")
library(tidyverse)

library(zoo)
library(dplyr)

library(vars)
```

```{r}
library(readr)
urlfile="https://raw.githubusercontent.com/Beau-Smit/Time-Series/main/data/MN_Ag_YieldsAndInputs.csv"
dat <- read.csv(url(urlfile))

#urlfile2="https://raw.githubusercontent.com/Beau-Smit/Time-Series/main/data/CMIP5_Agro.csv"
#dat2 <- read.csv(url(urlfile2), header = FALSE)
```

```{r}
#head(dat)
#tail(dat)
```

```{r}
# missing values
ag_df <- dat[!is.na(dat$Soy_BUperAcre),] %>% arrange(Year) 
#ag_df
```

- 3 useful variable: Corn_BUperAcre, Wheat_BUperAcre, Soy_BUperAcre, yearly data from 1934 to 2021
- each crop have 4 fertilizers, but too many nulls and gaps in those features ➔ will be ignored 

```{r}
# plot for relationship between variables
plot(ag_df$Corn_BUperAcre, ag_df$Wheat_BUperAcre, main = "Corn vs Wheat")
plot(ag_df$Corn_BUperAcre, ag_df$Soy_BUperAcre, main = "Corn vs Soy")
plot(ag_df$Soy_BUperAcre, ag_df$Wheat_BUperAcre, main = "Soy vs Wheat")
```

linear relationship

```{r}
# ts object
corn <- ts(ag_df$Corn_BUperAcre, frequency = 1, start = 1934)
wheat <- ts(ag_df$Wheat_BUperAcre, frequency = 1, start = 1934)
soy <- ts(ag_df$Soy_BUperAcre, frequency = 1, start = 1934)
```


```{r}
# corr
df <- cbind(corn, soy, wheat)
cor(df)
```


```{r}
# plot ts
autoplot(corn, main = "Crops Yield", xlab = "Year", ylab = "BU Per Acre", series = "Corn") +
  autolayer(wheat, series = "Wheat") +
  autolayer(soy, series = "Soy")
```

- boxcox needed
- potential seasonality or just trend? need to check

```{r}
# lambda calculation
BoxCox.lambda(corn)
BoxCox.lambda(soy)
BoxCox.lambda(wheat)
```

```{r}
# test with different lambda and plot ts to determine lambda for all three variables
corn_bc <- BoxCox(corn, lambda = 0.08)
soy_bc <- BoxCox(soy, lambda = 0.08)
wheat_bc <- BoxCox(wheat, lambda = 0.08)

autoplot(corn_bc, main = "Crops Yield after Boxcox", xlab = "Year", series = "Corn_bc") +
  autolayer(wheat_bc, series = "Wheat_bc") +
  autolayer(soy_bc, series = "Soy_bc")
```

can proceed with lambda = 0.5 (square root trans plus linear trans) 
or lambda = 0 (log trans) for easier interpretation

```{r}
# create variables for later use
corn_sqrt <- BoxCox(corn, lambda = 0.5)
soy_sqrt <- BoxCox(soy, lambda = 0.5)
wheat_sqrt <- BoxCox(wheat, lambda = 0.5)

corn_log <- BoxCox(corn, lambda = 0)
soy_log <- BoxCox(soy, lambda = 0)
wheat_log <- BoxCox(wheat, lambda = 0)

df_sqrt <- cbind(corn_sqrt, soy_sqrt, wheat_sqrt)
df_log <- cbind(corn_log, soy_log, wheat_log)

cor(df_sqrt) # check corr after boxcox
cor(df_log)
```


```{r}
# differencing and stationary check: determine d and D

diff(df_sqrt) %>% autoplot(main = "Crops Yield after sqrt+diff")
diff(df_log) %>% autoplot(maing = "Crops Yield after log+diff")

adf.test(diff(df_log)[,1])
adf.test(diff(df_sqrt)[,1])
```

both sqrt+1st differencing and log+1st differencing can bring to stationary ts, but log trans+1st differencing had relative constant variance.
- proceed with log + d=1 (no seasonal differencing)

```{r}
# acf/pacf to determine potential p,q,P,Q
# can use eacf for a mix of AR and MA
head(df_log)
diff(diff(df_log)[,3]) %>% acf()
diff(diff(df_log)[,3]) %>% pacf()

# nsdiffs(diff(df_log)[,1]): Non seasonal data -> P,Q = 0?
```

- no obvious seasonal component from ACF/PACF and nsdiffs() result
- corn can try (0,1,1) or (0,1,2)
- wheat can try (0,1,1)
- soy can try (0,1,1)

```{r}
# train/test split
split <- ts_split(df, sample.out = 8)

train <- split$train
test <- split$test

head(train)
```

#### Regression
```{r}
# regression
tslm_corn_log <- tslm(formula = corn ~ soy + wheat,
                      data = train,lambda = 0, biasadj = TRUE)
tslm_wheat_log <- tslm(formula = wheat ~ corn + soy,
                       data = train,lambda = 0, biasadj = TRUE)
tslm_soy_log <- tslm(formula = soy ~ corn + wheat,
                     data = train, lambda = 0, biasadj = TRUE)
```


```{r}
# plot regression fitted values & actual values in training set
autoplot(train[,1], series = "corn") + 
  autolayer(tslm_corn_log$fitted.values, series = "fitted corn")

autoplot(train[,2], series = "soy") +
  autolayer(tslm_soy_log$fitted.values, series = "fitted soy")

autoplot(train[,3], series = "wheat") +
  autolayer(tslm_wheat_log$fitted.values, series = "fitted wheat")

```


```{r}
#summary(tslm_corn_log)
```


```{r}
checkresiduals(tslm_corn_log)
checkresiduals(tslm_wheat_log)
checkresiduals(tslm_soy_log)
```

Breusch-Godfrey test for all three regression models had a low p-value, indicating residuals were autocorrelated ➔use regression with ARIMA error model is appropriate

#### Reg model with ARIMA errors
```{r}
# model estimation
regARIMA_corn_log <- auto.arima(train[,1], xreg = train[,2:3], d = 1, 
                                 lambda = 0, biasadj = TRUE, 
                                trace = TRUE, ic = "aicc")
  
regARIMA_wheat_log <- auto.arima(train[,3], xreg = train[,1:2], d = 1,
                                  lambda = 0, biasadj = TRUE, 
                                 trace = TRUE, ic = "aicc")
  
regARIMA_soy_log <- auto.arima(train[,2], xreg = train[,c(1,3)], d = 1,
                                lambda = 0, biasadj = TRUE, 
                               trace = TRUE, ic = "aicc")
```


```{r}
summary(regARIMA_corn_log)
summary(regARIMA_soy_log)
summary(regARIMA_wheat_log)
```


```{r}
# plot fitted values based on regression with ARIMA errors model vs actual values in training set
autoplot(train[,1], series = "corn") + 
  autolayer(regARIMA_corn_log$fitted, series = "fitted corn")

autoplot(train[,2], series = "soy") +
  autolayer(regARIMA_soy_log$fitted, series = "fitted soy")

autoplot(train[,3], series = "wheat") +
  autolayer(regARIMA_wheat_log$fitted, series = "fitted wheat")
```


```{r}
# residauls(), checkresiduals(), zero mean, (const variance)
# autocorr test
# normality test

checkresiduals(regARIMA_corn_log$residuals)
Box.test(regARIMA_corn_log$residuals, type =  "Ljung-Box")
shapiro.test(regARIMA_corn_log$residuals)

checkresiduals(regARIMA_soy_log$residuals)
Box.test(regARIMA_soy_log$residuals, type =  "Ljung-Box")
shapiro.test(regARIMA_soy_log$residuals)

checkresiduals(regARIMA_wheat_log$residuals)
Box.test(regARIMA_wheat_log$residuals, type =  "Ljung-Box")
shapiro.test(regARIMA_wheat_log$residuals)
```

Ljung-Box test had large p-value(> 0.05) for all three models, showing these model had captured patterns in the series so that there were no autocorrelation in the residuals

Shapiro-Wilk test had large p-value(>0.05) for all three models as well, indicating the residuals are normally distributed

```{r}
# forecast
head(test)

forecast_corn <- forecast(regARIMA_corn_log, h = 8, level = 95, xreg = test[,2:3])
forecast_soy <- forecast(regARIMA_soy_log, h = 8, level = 95, xreg = test[,c(1,3)])
forecast_wheat <- forecast(regARIMA_wheat_log, h = 8, level = 95, xreg = test[,1:2])
```


```{r}
# forecast values from the model vs actual values test setting 
autoplot(train[,1], main = "Corn Yield Forecast", xlab = "BU Per Acre", ylab = "Year") + autolayer(test[,1], series = "corn") + 
  autolayer(forecast_corn$mean, series = "forecasted corn")

autoplot(train[,2], main = "Soy Yield Forecast", xlab = "BU Per Acre", ylab = "Year") + autolayer(test[,2], series = "soy") +
  autolayer(forecast_soy$mean, series = "forecasted soy")

autoplot(train[,3], main = "Wheat Yield Forecast", xlab = "BU Per Acre", ylab = "Year") + autolayer(test[,3], series = "wheat") + autolayer(forecast_wheat$mean, series = "forecasted wheat")
```

```{r}
# model evaluation (testing set)
accuracy(forecast_corn, test[,1])
accuracy(forecast_soy, test[,2])
accuracy(forecast_wheat, test[,3])
```

```{r}
# test set accuracy computed with formula, ran and compare with above to make sure it's correct

## rmse
###corn
sqrt(mean((forecast_corn$mean - test[,1])^2))
###soy
###wheat

## mae
###corn
mean(abs(forecast_corn$mean - test[,1]))
###soy
###wheat

## mape
###corn
mean(abs((forecast_corn$mean - test[,1])*100/test[,1]))
###soy
###wheat
```


##### Cross Validation
```{r}
lead_years = 5 # fixed forecast length: 5 observations

rmse_df = data.frame()
mae_df = data.frame()
mape_df = data.frame()

for (idx in 40:(nrow(df) - 5)){ #40 is the fixed training set length for Sliding window, and 5 is the fixed forecast length for both sliding and expanding methods
  
  #iteration number
  it_number = idx - 39 #making 44 iteration total 
  
  test_cv = df[(idx + 1):(idx + lead_years), 1:3] #testing set for both methods
  
  train_expend = df[1:idx, 1:3] #training set for expanding method
  train_slide = df[(idx - 39):idx, 1:3] #training set for sliding method
  
  # Reg with ARIMA errors - Expanding
  modelCornExpand = auto.arima(train_expend[,1], xreg = train_expend[,2:3], d=1,
                               lambda = 0, biasadj = TRUE, ic = "aicc")
  modelWheatExpand = auto.arima(train_expend[,3], xreg = train_expend[,1:2], d=1,
                                lambda = 0, biasadj = TRUE, ic = "aicc")
  modelSoyExpand = auto.arima(train_expend[,2], xreg = train_expend[,c(1,3)], 
                              d=1, lambda = 0, biasadj = TRUE, ic = "aicc")
  
  forecastCornExpand = forecast(modelCornExpand, h = lead_years, level = 95, 
                               xreg = test_cv[,2:3])
  forecastWheatExpand = forecast(modelWheatExpand, h = lead_years, level = 95, 
                               xreg = test_cv[,1:2])
  forecastSoyExpand = forecast(modelSoyExpand, h = lead_years, level = 95, 
                               xreg = test_cv[,c(1,3)])
  
  mae_corn = mean(abs(forecastCornExpand$mean - test_cv[,1]))
  mae_wheat = mean(abs(forecastWheatExpand$mean - test_cv[,3]))
  mae_soy = mean(abs(forecastSoyExpand$mean - test_cv[,2]))
  
  rmse_corn = sqrt(mean((forecastCornExpand$mean - test_cv[,1])^2))
  rmse_wheat = sqrt(mean((forecastWheatExpand$mean - test_cv[,3])^2))
  rmse_soy = sqrt(mean((forecastSoyExpand$mean - test_cv[,2])^2))
  
  mape_corn = mean(abs((forecastCornExpand$mean - test_cv[,1])*100/test_cv[,1]))
  mape_wheat = mean(abs((forecastWheatExpand$mean - test_cv[,3])*100/test_cv[,3]))
  mape_soy = mean(abs((forecastWheatExpand$mean - test_cv[,2])*100/test_cv[,2]))
  
  newlinesMAE = data.frame(mae_corn, mae_wheat, mae_soy) %>%
    t() %>%
    cbind(data.frame(
      model = c("modelCornExpand", "modelWheatExpand", "modelSoyExpand"), 
      crop = c("Corn","Wheat","Soy"),
      n_train = nrow(train_expend)
      ))
  
  newlinesRMSE = data.frame(rmse_corn, rmse_wheat, rmse_soy) %>%
    t() %>%
    cbind(data.frame(
      model = c("modelCornExpand", "modelWheatExpand", "modelSoyExpand"), 
      crop = c("Corn","Wheat","Soy"),
      n_train = nrow(train_expend)
      ))
  
  newlinesMAPE = data.frame(mape_corn, mape_wheat, mape_soy) %>%
    t() %>%
    cbind(data.frame(
      model = c("modelCornExpand", "modelWheatExpand", "modelSoyExpand"), 
      crop = c("Corn","Wheat","Soy"),
      n_train = nrow(train_expend)
      ))
  
  mae_df = bind_rows(mae_df, newlinesMAE)
  rmse_df = bind_rows(rmse_df, newlinesRMSE)
  mape_df = bind_rows(mape_df, newlinesMAPE)  
  
  # Reg with ARIMA errors - Sliding
  modelCornSlide = auto.arima(train_slide[,1], xreg = train_slide[,2:3], d=1,
                               lambda = 0, biasadj = TRUE, ic = "aicc")
  modelWheatSlide = auto.arima(train_slide[,3], xreg = train_slide[,1:2], d=1,
                                lambda = 0, biasadj = TRUE, ic = "aicc")
  modelSoySlide = auto.arima(train_slide[,2], xreg = train_slide[,c(1,3)], 
                              d=1, lambda = 0, biasadj = TRUE, ic = "aicc")
  
  forecastCornSlide = forecast(modelCornSlide, h = lead_years, level = 95, 
                               xreg = test_cv[,2:3])
  forecastWheatSlide = forecast(modelWheatSlide, h = lead_years, level = 95, 
                               xreg = test_cv[,1:2])
  forecastSoySlide = forecast(modelSoySlide, h = lead_years, level = 95, 
                               xreg = test_cv[,c(1,3)])
  
  mae_corn = mean(abs(forecastCornSlide$mean - test_cv[,1]))
  mae_wheat = mean(abs(forecastWheatSlide$mean - test_cv[,3]))
  mae_soy = mean(abs(forecastSoySlide$mean - test_cv[,2]))  

  rmse_corn = sqrt(mean((forecastCornSlide$mean - test_cv[,1])^2))
  rmse_wheat = sqrt(mean((forecastWheatSlide$mean - test_cv[,3])^2))
  rmse_soy = sqrt(mean((forecastSoySlide$mean - test_cv[,2])^2))
  
  mape_corn = mean(abs((forecastCornSlide$mean - test_cv[,1])*100/test_cv[,1]))
  mape_wheat = mean(abs((forecastWheatSlide$mean - test_cv[,3])*100/test_cv[,3]))
  mape_soy = mean(abs((forecastSoySlide$mean - test_cv[,2])*100/test_cv[,2]))
  
  newlinesMAE = data.frame(mae_corn, mae_wheat, mae_soy) %>%
    t() %>%
    cbind(data.frame(
      model = c("modelCornSlide", "modelWheatSlide", "modelSoySlide"), 
      crop = c("Corn","Wheat","Soy"),
      n_train = nrow(train_slide)
      ))  
  
  newlinesRMSE = data.frame(rmse_corn, rmse_wheat, rmse_soy) %>%
    t() %>%
    cbind(data.frame(
      model = c("modelCornSlide", "modelWheatSlide", "modelSoySlide"), 
      crop = c("Corn","Wheat","Soy"),
      n_train = nrow(train_slide)
      ))
  
  newlinesMAPE = data.frame(mape_corn, mape_wheat, mape_soy) %>%
    t() %>%
    cbind(data.frame(
      model = c("modelCornSlide", "modelWheatSlide", "modelSoySlide"), 
      crop = c("Corn","Wheat","Soy"),
      n_train = nrow(train_slide)
      ))  

    
  mae_df = bind_rows(mae_df, newlinesMAE)
  rmse_df = bind_rows(rmse_df, newlinesRMSE)
  mape_df = bind_rows(mape_df, newlinesMAPE)  
}
```


```{r}
# DONT RUN THIS CHUNK!!
# just a coding test: one iteration only

lead_years = 5

rmse_df = data.frame()
mae_df = data.frame()
mape_df = data.frame()


  idx = 40
  
  test_cv = df[(idx + 1):(idx + lead_years), 1:3]
  
  train_expend = df[1:idx, 1:3]
  train_slide = df[(idx - 39):idx, 1:3]
  
  # Reg with ARIMA errors - Expanding
  modelCornExpand = auto.arima(train_expend[,1], xreg = train_expend[,2:3], d=1,
                               lambda = 0, biasadj = TRUE, ic = "aicc")
  modelWheatExpand = auto.arima(train_expend[,3], xreg = train_expend[,1:2], d=1,
                                lambda = 0, biasadj = TRUE, ic = "aicc")
  modelSoyExpand = auto.arima(train_expend[,2], xreg = train_expend[,c(1,3)], 
                              d=1, lambda = 0, biasadj = TRUE, ic = "aicc")
  
  forecastCornExpand = forecast(modelCornExpand, h = lead_years, level = 95, 
                               xreg = test_cv[,2:3])
  forecastWheatExpand = forecast(modelWheatExpand, h = lead_years, level = 95, 
                               xreg = test_cv[,1:2])
  forecastSoyExpand = forecast(modelSoyExpand, h = lead_years, level = 95, 
                               xreg = test_cv[,c(1,3)])
  
  mae_corn = mean(abs(forecastCornExpand$mean - test_cv[,1]))
  mae_wheat = mean(abs(forecastWheatExpand$mean - test_cv[,3]))
  mae_soy = mean(abs(forecastSoyExpand$mean - test_cv[,2]))
  
  rmse_corn = sqrt(mean((forecastCornExpand$mean - test_cv[,1])^2))
  rmse_wheat = sqrt(mean((forecastWheatExpand$mean - test_cv[,3])^2))
  rmse_soy = sqrt(mean((forecastSoyExpand$mean - test_cv[,2])^2))
  
  mape_corn = mean(abs((forecastCornExpand$mean - test_cv[,1])*100/test_cv[,1]))
  mape_wheat = mean(abs((forecastWheatExpand$mean - test_cv[,3])*100/test_cv[,3]))
  mape_soy = mean(abs((forecastSoyExpand$mean - test_cv[,2])*100/test_cv[,2]))
  
  newlinesMAE = data.frame(mae_corn, mae_wheat, mae_soy) %>%
    t() %>%
    cbind(data.frame(
      model = c("modelCornExpand", "modelWheatExpand", "modelSoyExpand"), 
      crop = c("Corn","Wheat","Soy"),
      n_train = nrow(train_expend)
      ))
  
  newlinesRMSE = data.frame(rmse_corn, rmse_wheat, rmse_soy) %>%
    t() %>%
    cbind(data.frame(
      model = c("modelCornExpand", "modelWheatExpand", "modelSoyExpand"), 
      crop = c("Corn","Wheat","Soy"),
      n_train = nrow(train_expend)
      ))
  
  newlinesMAPE = data.frame(mape_corn, mape_wheat, mape_soy) %>%
    t() %>%
    cbind(data.frame(
      model = c("modelCornExpand", "modelWheatExpand", "modelSoyExpand"), 
      crop = c("Corn","Wheat","Soy"),
      n_train = nrow(train_expend)
      ))
  
  mae_df = bind_rows(mae_df, newlinesMAE)
  rmse_df = bind_rows(rmse_df, newlinesRMSE)
  mape_df = bind_rows(mape_df, newlinesMAPE)
  
  # Reg with ARIMA errors - Sliding
  modelCornSlide = auto.arima(train_slide[,1], xreg = train_slide[,2:3], d=1,
                               lambda = 0, biasadj = TRUE, ic = "aicc")
  modelWheatSlide = auto.arima(train_slide[,3], xreg = train_slide[,1:2], d=1,
                                lambda = 0, biasadj = TRUE, ic = "aicc")
  modelSoySlide = auto.arima(train_slide[,2], xreg = train_slide[,c(1,3)], 
                              d=1, lambda = 0, biasadj = TRUE, ic = "aicc")
  
  forecastCornSlide = forecast(modelCornSlide, h = lead_years, level = 95, 
                               xreg = test_cv[,2:3])
  forecastWheatSlide = forecast(modelWheatSlide, h = lead_years, level = 95, 
                               xreg = test_cv[,1:2])
  forecastSoySlide = forecast(modelSoySlide, h = lead_years, level = 95, 
                               xreg = test_cv[,c(1,3)])
  
  mae_corn = mean(abs(forecastCornSlide$mean - test_cv[,1]))
  mae_wheat = mean(abs(forecastWheatSlide$mean - test_cv[,3]))
  mae_soy = mean(abs(forecastSoySlide$mean - test_cv[,2]))  
  
  rmse_corn = sqrt(mean((forecastCornSlide$mean - test_cv[,1])^2))
  rmse_wheat = sqrt(mean((forecastWheatSlide$mean - test_cv[,3])^2))
  rmse_soy = sqrt(mean((forecastSoySlide$mean - test_cv[,2])^2))
  
  mape_corn = mean(abs((forecastCornSlide$mean - test_cv[,1])*100/test_cv[,1]))
  mape_wheat = mean(abs((forecastWheatSlide$mean - test_cv[,3])*100/test_cv[,3]))
  mape_soy = mean(abs((forecastSoySlide$mean - test_cv[,2])*100/test_cv[,2]))
  
  newlinesMAE = data.frame(mae_corn, mae_wheat, mae_soy) %>%
    t() %>%
    cbind(data.frame(
      model = c("modelCornSlide", "modelWheatSlide", "modelSoySlide"), 
      crop = c("Corn","Wheat","Soy"),
      n_train = nrow(train_slide)
      ))
  
  newlinesRMSE = data.frame(rmse_corn, rmse_wheat, rmse_soy) %>%
    t() %>%
    cbind(data.frame(
      model = c("modelCornSlide", "modelWheatSlide", "modelSoySlide"), 
      crop = c("Corn","Wheat","Soy"),
      n_train = nrow(train_slide)
      ))
  
  newlinesMAPE = data.frame(mape_corn, mape_wheat, mape_soy) %>%
    t() %>%
    cbind(data.frame(
      model = c("modelCornSlide", "modelWheatSlide", "modelSoySlide"), 
      crop = c("Corn","Wheat","Soy"),
      n_train = nrow(train_slide)
      ))
  
  mae_df = bind_rows(mae_df, newlinesMAE)
  rmse_df = bind_rows(rmse_df, newlinesRMSE)
  mape_df = bind_rows(mape_df, newlinesMAPE)
```


```{r}
rownames(mae_df) = 1:nrow(mae_df)
names(mae_df)[1] <- "MAE"
mae_df

rownames(rmse_df) = 1:nrow(rmse_df)
names(rmse_df)[1] <- "RMSE"
rmse_df

rownames(mape_df) = 1:nrow(mape_df)
names(mape_df)[1] <- "MAPE"
mape_df
```


```{r}

#Corn accuracy simply from the Avg of all 44 iteration
mae <- split(mae_df, mae_df$crop)
mean(mae$Corn$MAE)

rmse <- split(rmse_df, rmse_df$crop)
mean(rmse$Corn$RMSE)

mape <- split(mape_df, mape_df$crop)
mean(mape$Corn$MAPE)

```


```{r}

#Wheat accuracy simply from the Avg of all 44 iteration

mean(mae$Wheat$MAE)

mean(rmse$Wheat$RMSE)

mean(mape$Wheat$MAPE)

```


```{r}

#Soy accuracy simply from the Avg of all 44 iteration

mean(mae$Soy$MAE)

mean(rmse$Soy$RMSE)

mean(mape$Soy$MAPE)

```