---
title: "Ag Yield Prediction - VAR"
output: html_document
---

```{r setup, include=FALSE}
library(forecast)
library(knitr)
library(vars)
library(tidyverse)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE)
```

Step 0) read in Corn, Soy, and Wheat data for Minnesota:

```{r}
ag_df_all = read_csv("../data/MN_Ag_YieldsAndInputs.csv") %>% arrange(Year)
bedd_and_precip = read_csv("../data/CMIP5_Agro_year.csv", col_names = FALSE)
colnames(bedd_and_precip) = c('idx', 'value', 'year', 'variable', 'scenario')
bedd_and_precip = bedd_and_precip %>%
  pivot_wider(names_from = variable, values_from = value) %>% 
  dplyr::select(year, BEDD, RR1, scenario)

```

For the purposes of the VAR, we will drop the rows that don't have a value for all three crops. Ultimately, this is observations rows (1934 - 2021).

```{r}
ag_df = ag_df_all[!is.na(ag_df_all$Soy_BUperAcre),]
head(ag_df)
```
As we can see here, we don't have fertilizer input information for each year of data. We can use a spline approach to interpolate the missing values, but this will be trickier for the newer values as we don't have data on either side and we're essentially forecasting things that already happened.

```{r}
ag_plot = ag_df %>% pivot_longer(cols = ends_with('BUperAcre'))
ggplot(ag_plot, aes(x = Year, y = value, col = name)) + 
    geom_line() +
    ggtitle('Yield - BU Per Acre') +
    scale_color_discrete(name = 'Crop', labels = c('Corn', 'Soybeans', 'Wheat')) + 
    ylab('BU / Acre')
```

Our data look a bit heteroskedastic - especially in the down years. Let's use a box cox transformation and see if we can make an improvement.

```{r}
ag_yields = ag_df[,1:4]
BoxCox.lambda(ts(ag_yields[,2:4]))
```
This is pretty close to a straight log transformation.

```{r}
ag_plot = ag_df[,2:4] %>%
    BoxCox(0.04355743) %>%
    cbind(Year = ag_df[,1]) %>%
    pivot_longer(cols = ends_with('BUperAcre'))

ggplot(ag_plot, aes(x = Year, y = value, col = name)) + 
    geom_line() +
    ggtitle('Yield - BU Per Acre (Box Cox Transformed with lambda = .044)') +
    scale_color_discrete(name = 'Crop', labels = c('Corn', 'Soybeans', 'Wheat')) + 
    ylab('BU / Acre')
```

This makes the perturbations look a bit less extreme - we will go with this moving forward. 

```{r}
# ag_lambda = 0.04355743
# 
# ag_transformed = ag_df[,2:4] %>%
#     BoxCox(0.04355743) %>%
#     cbind(Year = ag_df[,1])
# 
# lead_years = 5
# mae_df = data.frame()
# aicc_df = data.frame()
# 
# for (p_lag in 5:1) {
# 
#     for (idx in 40:(nrow(ag_transformed) - 1)){
#       #increase iteration number
#       it_number = idx - 39
#     
#       test_ts = ag_transformed[(idx + 1):(idx + lead_years), 1:3]
#       
#       train_expand = ag_transformed[1:idx, 1:3] %>% ts()
#       train_slide = ag_transformed[(idx - 39):idx, 1:3] %>% ts()
#       
#       # VAR Expanding
#       m1 = VAR(train_expand, p = p_lag, type = 'both')
#       f1 = forecast(m1, h=lead_years)
# 
#       mae_corn = abs(f1$forecast$Corn_BUperAcre$mean - test_ts$Corn_BUperAcre)
#       mae_wheat = abs(f1$forecast$Wheat_BUperAcre$mean - test_ts$Wheat_BUperAcre)
#       mae_soy = abs(f1$forecast$Soy_BUperAcre$mean - test_ts$Soy_BUperAcre)
#     
#       newlines = data.frame(mae_corn, mae_wheat, mae_soy) %>%
#           t() %>% 
#           cbind(data.frame(model = 'Expanding VAR',
#                            crop = c('Corn', 'Wheat', 'Soy'),
#                            n_train = nrow(train_expand),
#                            lag_order = p_lag))
# 
#       mae_df = bind_rows(mae_df, newlines)
# 
#       # VAR Sliding
#       m2 = VAR(train_slide, p = p_lag, type = 'both')
#       f2 = forecast(m2, h=lead_years)
#       
#       mae_corn = abs(f2$forecast$Corn_BUperAcre$mean - test_ts$Corn_BUperAcre)
#       mae_wheat = abs(f2$forecast$Wheat_BUperAcre$mean - test_ts$Wheat_BUperAcre)
#       mae_soy = abs(f2$forecast$Soy_BUperAcre$mean - test_ts$Soy_BUperAcre)
#     
#       newlines = data.frame(mae_corn, mae_wheat, mae_soy) %>%
#           t() %>% 
#           cbind(data.frame(model = 'Sliding VAR',
#                            crop = c('Corn', 'Wheat', 'Soy'),
#                            n_train = nrow(train_slide),
#                            lag_order = p_lag))
#       
#       mae_df = bind_rows(mae_df, newlines)
#     }    
# }

```

```{r}
# rownames(mae_df) = 1:nrow(mae_df)
# plot_mae = pivot_longer(mae_df, cols = -c('model', 'crop', 'lag_order', 'n_train')) %>%
#     rename(horizon = name,
#            abs_error = value) %>% 
#     mutate(horizon = as.integer(horizon))
```

```{r fig.width=10, fig.height=14}
# ggplot(plot_mae, aes(x = horizon, y = abs_error, col=model)) +
#   facet_grid(paste0('lag ', lag_order)~crop) + 
#   geom_smooth() + 
#   ylab('MAE')
```

## Adding in Agroclimatic CMIP5 indicators:

```{r}
bedd_and_precip = read_csv("../data/CMIP5_Agro_year.csv", col_names = FALSE)
colnames(bedd_and_precip) = c('idx', 'value', 'year', 'variable', 'scenario')
bedd_and_precip = bedd_and_precip %>%
  pivot_wider(names_from = variable, values_from = value) %>% 
  dplyr::select(year, BEDD, RR1, scenario)
head(bedd_and_precip)
```

```{r}
tsdisplay(ag_df_exog$BEDD)
```
```{r}
tsdisplay(ag_df_exog$RR1)
```

```{r}
hist_agro = bedd_and_precip %>% filter(scenario == 'hist') %>% rename(Year = year)
ag_df_exog = ag_df_all %>% left_join(hist_agro) %>% 
  filter(!is.na(BEDD))
```
```{r}
BoxCox.lambda(ag_df_exog[,c('BEDD', 'RR1')])
```


```{r}
ag_lambda = 0

ag_transformed = ag_df_exog[,2:4] %>%
    BoxCox(0) %>%
    cbind(Year = ag_df_exog[,'Year'],
          BEDD = ag_df_exog[,'BEDD'],
          RR1 = ag_df_exog[,'RR1'])

ts_yield = ag_transformed[, 1:3]
ts_exog = ag_transformed[, c('BEDD','RR1')]
VARselect(ts_yield, exogen = ts_exog, lag.max = p_lag, type = 'both')
```
Great news - let's see if it passes the Portmanteau test. We would love to _accept_ the null hypothesis of no serial correlation amongst residuals (meaning there is some autocorrelation left on the table for a different model to make use of).

```{r}
VAR_best = VAR(ts_yield, exogen = ts_exog, p = 1, type = "both")

serial.test(VAR_best, type="PT.asymptotic")
```
Great success - we accept the null hypothesis of no serial correlation among residuals. Given the unanimous best VAR model having only one lag (already very parsimonious) we will crossvalidate with this moving forward.

## Comparison with other models

- 5 leads 
- 1 lag
- 60 year sliding window
- BoxCox lambda of 0 - log transformation, as it's pretty close to 0 anyways and it helps with interpretability
- mean percent error for each lead to facilitate comparison between different crops


### First, straight VAR with no exogenous predictors, with all three yields.
```{r}
ag_lambda = 0

ag_transformed = ag_df[,2:4] %>%
    BoxCox(0) %>%
    cbind(Year = ag_df[,1])

lead_years = 5
window_len = 60
p_lag = 5
ape_df = data.frame()

for (idx in window_len:(nrow(ag_transformed) - 1)){
    test_ts = ag_transformed[(idx + 1):(idx + lead_years), 1:3]
    train_slide = ag_transformed[(idx - window_len):idx, 1:3] %>% ts()
  
    # VAR Sliding
    m2 = VAR(train_slide, p = 1, type = 'both')
    f2 = forecast(m2, h=lead_years)
    year = max(ag_transformed[(idx - window_len):idx, 'Year'])

    ape_corn = abs(f2$forecast$Corn_BUperAcre$mean - test_ts$Corn_BUperAcre) / test_ts$Corn_BUperAcre
    ape_wheat = abs(f2$forecast$Wheat_BUperAcre$mean - test_ts$Wheat_BUperAcre) / test_ts$Wheat_BUperAcre
    ape_soy = abs(f2$forecast$Soy_BUperAcre$mean - test_ts$Soy_BUperAcre) / test_ts$Soy_BUperAcre
  
    forecasts = data.frame(f2$forecast$Corn_BUperAcre$mean,
                           f2$forecast$Wheat_BUperAcre$mean,
                           f2$forecast$Soy_BUperAcre$mean) %>% t()

    colnames(forecasts) = paste('VAR3_t+', 1:5, sep='')

    newlines = data.frame(ape_corn, ape_wheat, ape_soy, row.names = paste('APE_', 1:5, sep='')) %>%
        t() %>% 
        cbind(forecasts) %>% 
        cbind(data.frame(model = 'VAR3',
                         crop = c('corn', 'wheat', 'soy'),
                         year = year))
    
    ape_df = bind_rows(ape_df, newlines)
}
rownames(ape_df) = 1:nrow(ape_df)
kable(ape_df, digits = 3)
```

```{r}
plot_ape = pivot_longer(ape_df, cols = -c('model', 'crop', paste('VAR3_t+', 1:5, sep=''))) %>%
    rename(horizon = name,
           abs_error = value) %>% 
    mutate(horizon = as.integer(str_sub(horizon, -1)))

ggplot(plot_ape, aes(x = horizon, y = abs_error, col=model)) +
  facet_grid(.~crop) + 
  geom_smooth() + 
  ylab('MAPE')
```

# Next, VAR(1) with just Wheat and Corn
```{r}
ag_lambda = 0
ag_df = ag_df_all[!is.na(ag_df_all$Wheat_BUperAcre),]
  
ag_transformed = ag_df[,2:4] %>%
    BoxCox(0) %>%
    cbind(Year = ag_df[,1])

lead_years = 5
window_len = 60
p_lag = 5
ape_df2 = data.frame()

for (idx in window_len:(nrow(ag_transformed) - 1)){
    test_ts = ag_transformed[(idx + 1):(idx + lead_years), 1:2]
    train_slide = ag_transformed[(idx - window_len):idx, 1:2] %>% ts()
    year = max(ag_transformed[(idx - window_len):idx, 'Year'])

    # VAR Sliding
    m2 = VAR(train_slide, p = 1, type = 'both')
    f2 = forecast(m2, h=lead_years)
    
    ape_corn = abs(f2$forecast$Corn_BUperAcre$mean - test_ts$Corn_BUperAcre) / test_ts$Corn_BUperAcre
    ape_wheat = abs(f2$forecast$Wheat_BUperAcre$mean - test_ts$Wheat_BUperAcre) / test_ts$Wheat_BUperAcre
    
    forecasts = data.frame(f2$forecast$Corn_BUperAcre$mean,
                           f2$forecast$Wheat_BUperAcre$mean) %>% t()

    colnames(forecasts) = paste('VARcw_t+', 1:5, sep='')

    newlines = data.frame(ape_corn, ape_wheat, row.names = paste('APE_', 1:5, sep='')) %>%
        t() %>% 
        cbind(forecasts) %>% 
        cbind(data.frame(model = 'VARcw',
                         crop = c('corn', 'wheat'),
                         year = year))
    
    ape_df2 = bind_rows(ape_df2, newlines)
}
rownames(ape_df2) = 1:nrow(ape_df2)
kable(ape_df2, digits = 3)
```

```{r}
plot_ape = pivot_longer(ape_df2, cols = -c('model', 'crop', paste('VARcw_t+', 1:5, sep=''))) %>%
    rename(horizon = name,
           abs_error = value) %>% 
    mutate(horizon = as.integer(str_sub(horizon, -1)))

ggplot(plot_ape, aes(x = horizon, y = abs_error, col=model)) +
  facet_grid(.~crop) + 
  geom_smooth() + 
  ylab('MAPE')
```


### Last, VAR(1) with all three crops AND CMIP5 historical predictors. Due to to the small sample size (59 years) we decrease window length to 30.
```{r}
ag_df = ag_df_exog[!is.na(ag_df_exog$RR1),]

ag_transformed = ag_df[,2:4] %>%
    BoxCox(0) %>%
    cbind(Year = ag_df[,'Year'],
          BEDD = ag_df[,'BEDD'],
          RR1 = ag_df[,'RR1'])

lead_years = 5
window_len = 30
p_lag = 5
ape_df3 = data.frame()
VARspec = data.frame()
for (idx in window_len:(nrow(ag_transformed) - 1)){

    test_ts = ag_transformed[(idx + 1):(idx + lead_years), 1:3] 
    exog_test_ts = ag_transformed[(idx + 1):(idx + lead_years), c('BEDD','RR1')] 
    train_slide = ag_transformed[(idx - window_len):idx, 1:3] %>% ts()
    exog_slide = ag_transformed[(idx - window_len):idx, c('BEDD','RR1')] %>% ts()

    # VAR Sliding
    m2 = VAR(train_slide, exogen = exog_slide, p=1, type = 'both')
    f2 = forecast(m2, h=lead_years, dumvar = exog_test_ts)
    year = max(ag_transformed[(idx - window_len):idx, 'Year'])

    ape_corn = abs(f2$forecast$Corn_BUperAcre$mean - test_ts$Corn_BUperAcre) / test_ts$Corn_BUperAcre
    ape_wheat = abs(f2$forecast$Wheat_BUperAcre$mean - test_ts$Wheat_BUperAcre) / test_ts$Wheat_BUperAcre
    ape_soy = abs(f2$forecast$Soy_BUperAcre$mean - test_ts$Soy_BUperAcre) / test_ts$Soy_BUperAcre

    forecasts = data.frame(f2$forecast$Corn_BUperAcre$mean,
                           f2$forecast$Wheat_BUperAcre$mean,
                           f2$forecast$Soy_BUperAcre$mean) %>% t()

    colnames(forecasts) = paste('VAR3exog_t+', 1:5, sep='')

    newlines = data.frame(ape_corn, ape_wheat, ape_soy, row.names = paste('APE_', 1:5, sep='')) %>%
        t() %>%
        cbind(forecasts) %>%
        cbind(data.frame(model = 'VAR3exog',
                         crop = c('corn', 'wheat', 'soy'),
                         year = year))

    ape_df3 = bind_rows(ape_df3, newlines)
}
rownames(ape_df3) = 1:nrow(ape_df3)
kable(ape_df3, digits=2)
```



```{r}
df_out = ape_df2 %>% dplyr::select(starts_with('VARcw_t+'), year, crop) %>% 
  full_join(ape_df3 %>% dplyr::select(starts_with('VAR3exog_t+'), year, crop)) %>%
  full_join(ape_df %>% dplyr::select(starts_with('VAR3_t+'), year, crop)) %>%
  mutate_at(vars(starts_with('VAR')), exp) %>% 
  arrange(crop)

df_out %>% write.csv('../data/VAR_predictions.csv')
```

